Collaborative Multioutput Gaussian Processes

### 2014-03-15 paper/experiments.tex
latest experiment results and paper style

%todo: result for independent gps
We also report in Table \ref{tab:fx} the predictive accuracy of our model compared to the convolved GPs model with exact inference \citet{alvarez-lawrence-nips-08} (CGP) and with approximation via the variational inducing kernels \citet{alvarez2010efficient}) (CGPVAR) in addition to using independent GPs (IGP, one for each output).
Our model outperforms both of the CGP variants in terms of standardized mean-squared error (SMSE).
It has lower test likelihood, as measured by the negative log predictive density (NLPD), which is mainly due to the less conservative predictive variance of the exact CGP for the CAD currency.
The results are averaged across the 3 outputs over 5 repetitions.
% no std because there are 3 outputs but variance is small
Note that the SMSE of CGPVAR is taken from \citet{alvarez2010efficient} while the NLPD was not provided.
Training took only 10 minutes for our model compared to 1.4 hours of the full CGP model.

\begin{table}[h]
\caption{Performance comparison on the foreign exchange rate dataset.}
\label{tab:fx}
\begin{center}
\begin{tabular}{c|c|c}
METHOD & SMSE & NLPD \\ \hline
COGP  & \textbf{0.2125} & -0.8394 \\
CGP & 0.2427 & \textbf{-2.9474} \\
IGP & 0.5996 & 0.4082 \\
CGPVAR & 0.2795 & NA 
%ICM & 0.3927 &
\end{tabular}
\end{center}
\end{table}

@@

The real data and the predictive distributions by 3 models (COGP, CGP with exact inference, and independent GPs) are shown in Figure \ref{fig:weather}.
It is clear that the independent GPs model is clueless in the test regions and thus simply uses the average temperature as its prediction.
For Cambermet, both COGP and CGP can capture the rising in temperature from the morning till the afternoon and the fall afterward.
For Chimet, both models perform poorly but CGP more so as it falsely predicts  wild fluctuations that are non-existent in the data. 
Note that the two latent processes learned by our model indeed correspond to different patterns in the data: one process has the inverse lengthscale of 136 which captures the global increase in temperature during the training period while the other has the inverse lengthscale of 0.5 to model the local variations within a single day.
The comparative performance of the models are summarized in Table \ref{tab:air}, which further supports the qualitative analysis that our model outperforms CGP.
All results are averaged of the 2 outputs over 5 repetitions.
Training of our model took on average 5 minutes compared to 3 hours of CGP with exact inference.

\begin{table}[h]
\caption{Performance comparison on the air temperature dataset.}
\label{tab:air}
\begin{center}
\begin{tabular}{c|c|c}
METHOD & SMSE & NLPD \\ \hline
COGP & \textbf{0.1077} & \textbf{2.1712} \\
CGP & 0.1125 & 2.2219 \\
IGP & 0.8944 & 12.5319
\end{tabular}
\end{center}
\end{table}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/slfm-weatherCambermet.eps} &
\includegraphics[scale=0.3]{figures/cgp-weatherCambermet.eps} &
\includegraphics[scale=0.3]{figures/weatherCambermet.eps}
\\
\includegraphics[scale=0.3]{figures/slfm-weatherChimet.eps} &
\includegraphics[scale=0.3]{figures/cgp-weatherChimet.eps} &
\includegraphics[scale=0.3]{figures/weatherChimet.eps}
\end{tabular}
\caption{Predictive distributions of the multioutput GPs (left column) and full independent GPs (right column) for the air temperature problem.}
\label{fig:weather}
\end{figure*}

@@

>>>>

%todo: result for independent gps
We also report in Table \ref{tab:fx} the predictive accuracy of our model compared to the convolved GPs model with exact inference \citet{alvarez-lawrence-nips-08} (CGP) and with approximation via the variational inducing kernels \citet{alvarez2010efficient}) (CGPVAR) in addition to using independent GPs (IGP, one for each output).
Our model outperforms both of the CGP variants in terms of standardized mean-squared error (SMSE).
It has lower test likelihood, as measured by the negative log predictive density (NLPD), which is mainly due to the less conservative predictive variance of the exact CGP for the CAD currency.
%The results are averaged across the 3 outputs over 5 repetitions.
% no std because there are 3 outputs but variance is small
Note that the SMSE of CGPVAR is taken from \citet{alvarez2010efficient} while the NLPD was not provided.
Training took only 10 minutes for our model compared to 1.4 hours of the full CGP model.

\setlength{\tabcolsep}{4pt}
\begin{table}[t]
\caption{Performance comparison on the foreign exchange rate dataset. Results are averaged of the 3 outputs over 5 repetitions.}
\label{tab:fx}
\begin{center}
\begin{tabular}{ccc}
\toprule
\textbf{METHOD} & \textbf{SMSE} & \textbf{NLPD} \\ \hline
COGP  & \textbf{0.2125} & -0.8394 \\
CGP & 0.2427 & \textbf{-2.9474} \\
IGP & 0.5996 & 0.4082 \\
CGPVAR & 0.2795 & NA \\
%ICM & 0.3927 &
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

The real data and the predictive distributions by 3 models (COGP, CGP with exact inference, and independent GPs) are shown in Figure \ref{fig:weather}.
It is clear that the independent GPs model is clueless in the test regions and thus simply uses the average temperature as its prediction.
For Cambermet, both COGP and CGP can capture the rising in temperature from the morning till the afternoon and the fall afterward.
For Chimet, both models perform poorly but CGP more so as it falsely predicts  wild fluctuations that are non-existent in the data. 
Note that the two latent processes learned by our model indeed correspond to different patterns in the data: one process has the inverse lengthscale of 136 which captures the global increase in temperature during the training period while the other has the inverse lengthscale of 0.5 to model the local variations within a single day.
The comparative performance of the models are summarized in Table \ref{tab:air}, which further supports the qualitative analysis that our model outperforms CGP.
%All results are averaged of the 2 outputs over 5 repetitions.
Training of our model took on average 5 minutes compared to 3 hours of CGP with exact inference.

\setlength{\tabcolsep}{4pt}
\begin{table}[t]
\caption{Performance comparison on the air temperature dataset. Results are averaged of 2 outputs over 5 repetitions. }
\label{tab:air}
\begin{center}
\begin{tabular}{ccc}
\toprule
\textbf{METHOD} & \textbf{SMSE} & \textbf{NLPD} \\
\hline
COGP & \textbf{0.1077} & \textbf{2.1712} \\
CGP & 0.1125 & 2.2219 \\
IGP & 0.8944 & 12.5319 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/cogp-weatherCambermet.eps} &
\includegraphics[scale=0.3]{figures/cgp-weatherCambermet.eps} &
\includegraphics[scale=0.3]{figures/weatherCambermet.eps}
\\
\includegraphics[scale=0.3]{figures/cogp-weatherChimet.eps} &
\includegraphics[scale=0.3]{figures/cgp-weatherChimet.eps} &
\includegraphics[scale=0.3]{figures/weatherChimet.eps}
\end{tabular}
\caption{Predictive distributions of the multioutput GPs (left column) and full independent GPs (right column) for the air temperature problem.}
\label{fig:weather}
\end{figure*}

@@

Since none of the existing multioutput models are applicable to this problem, we compare with two baselines which learn each output independently.
The first is the subset of data approach where a subset of 1000 data points (for each output) is randomly sampled for training. 
We call this method SOD1000.
The second baseline is the sparse GP with stochastic variational inference \citet{hensmangaussian} using 500 inducing inputs, which we refer to as SVIGP.
The outputs are normalized and the squared exponential with ARD covariance function is used for all methods.
In case of COGP, this mean the shared process (we used $Q = 1$) and each of the independent processes have this same covariance function.

The predictive performance of the methods are given in Table \ref{tab:robotarm}.
The results are averaged of the 2 outputs over 5 repetitions.
Both COGP and SVIGP using 500 inducing inputs significantly outperform SOD in terms of the mean absolute error (MAE) but have lower predictive likelihood.

\begin{table}[t]
\caption{Performance comparison on the robot arm dataset. In the two bottom lines, standard GP is applied to output 1 and the remaining method is applied to output 2. Results are averaged over 5 repetitions.}
\label{tab:robotarm}
\begin{center}
\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{\textbf{OUTPUT 1}} & \multicolumn{2}{c}{\textbf{OUTPUT 2}} \\ \cmidrule(r){2-5}
\textbf{METHOD} & \textbf{SMSE} & \textbf{NLPD} & \textbf{SMSE} & \textbf{NLPD}\\ 
 \midrule
COGP, learn z & \textbf{0.2631} & \textbf{3.0600} & 0.0127 & \textbf{0.8302} \\
COGP, fix z & 0.2821& 3.2281 & 0.0131 & 0.8685 \\
GP, SVIGP & 0.3119 & 3.2198 & \textbf{0.0101} & 1.1914 \\
GP, SOD & 0.3119 & 3.2198 & 0.0104 & 1.9407 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\subsubsection{Impact of Optimizing the Inducing Inputs}
% table for smses, nlpds, training tiem
\begin{figure}
\includegraphics[scale=0.4]{figures/sarcosTime.eps}
\caption{The cost of optimizing versus fixing the inducing inputs.}
\label{fig:time}
\end{figure}

### 2014-03-15 paper/paper.tex
updated paper
clean figures
updated figures and description of experiment on the fx dataset
move appendix to supplemental and re-format
latest experiment results and paper style

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{latexsym,amsbsy,amssymb,color,xspace, booktabs}

@@

\title{Stochastic Multiple Output Gaussian Processes}

@@

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). Reference style should follow the style
that you are used to using, as long as the citation style is
consistent.

@@

\section{DISCUSSION}
\input{discussion}
\newpage
\section{APPENDIX}
Perhaps this should go to a supplementary material.
\input{appendix}

@@

\bibliographystyle{abbrv}
\bibliography{references}

>>>>

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{latexsym,amsbsy,amssymb,color,xspace,booktabs}

@@

\title{Collaborative Multioutput Gaussian Processes}

@@

@@

\section{DISCUSSION}
\input{discussion}

@@

\bibliographystyle{plainnat}
\bibliography{references}
